---
title: "Data Manipulation Assignment"
author: "Cali Keller"
date: "2025-10-09"
output: html_document
---

#Part 1
1.Open the sleep dataset in R 
```{r}
data("sleep")
library(tidyr)
library(tidyverse)
```
√; tidyverse and tidyr are redundant
2. Is it long or wide format? How do you know? 
```{r}
#The data is in long format, as there are two repeated groups.
```
√
3. Transform this data so that it is in the other format.
```{r}
sleepwide <- pivot_wider(sleep, id_cols = "ID", names_from = "group", values_from = "extra")
```
√
4. Transform your data back to its original format. 
```{r}
sleeplong <- pivot_longer(sleepwide, cols = c("1", "2"), names_to = "extra", values_to = "group")
```
√
5. Open the Starwars dataset
```{r}
data("starwars")
```
√
6. Create homeworld as a factor. How many levels does it have? 
```{r}
homeworldfactor <- as.factor(starwars$homeworld)
#it has 48 levels
```
√
7. Reorder homeworld so that Naboo is the reference group. 
```{r}
homeworldfactor == "Naboo"
```
this doesn't work. need to do relevel() and use ref="Naboo" 
8. Use mutate to create a column transforming kg to pounds. (hint: you may need to look up the conversion)
```{r}
starwars %>% mutate(kg = mass * 2.2046) 
```
√
9. Print a summary of your column from #8
```{r}
summary(starwars$mass)
```
√

#Part 2
10. open the billboard dataset.
```{r}
data(billboard)
```
√
11. are the data in this dataframe in long or wide format? explain how you know.
```{r}
#this is in wide format 
```
√
12. change the data from its original format to the other version. 
```{r}
billboardlong <- pivot_longer(billboard, cols = 4:79, names_to = "week", values_to = "score" )
```
√
13. change it from your altered version in #3 back to its original state. 
```{r}
billboardwide <- pivot_wider(billboardlong, names_from = "week", values_from = "score" )
```
√
14. Separate the date entered column into three columns: one for year, one for month, and one for day. 
```{r}
separate(billboardwide, date.entered, into = c("year", "month", "day"), sep = "-")
```
√
