---
title: "foundations in r pt 2 hw"
author: "Cali Keller"
date: "2025-10-02"
output: html_document
---
1. Load the tidyverse library. 
```{r}
library(tidyverse)
```

2. Load the billboard data. 
```{r}
data(billboard)
```

3. What type of data is this? (e.g. describe what's in the dataset).
```{r}
#Song rankings for Billboard top 100 in the year 2000

```

4. Print the column names from this dataset. 
```{r}
colnames(billboard)
```

5. Print the first 5 entries in the dataset. 
```{r}
billboard[1:5, ]
```

# Practice
1. Create a subset of data that has Jay-Z as the only artist. How many songs are in it?
```{r}
jayz <- subset(billboard, billboard$artist == "Jay-Z")
summary(jayz)
# there are five songs in this subset 
```

2. In the original dataset, use filter to find information about the song "All the small things." What was the highest position (i.e. lowest number) this song reached on the billboard 100? 
```{r}
filter(billboard, billboard$track == "All The Small Things")
#it got to number 6.
```

3. Create a subset of data for songs Tim McGraw sang. Summarize the data for wk 1.
```{r}
timmy <- subset(billboard, billboard$artist == "McGraw, Tim")
summary(timmy$wk1)
```

4. Create a function that can be used to determine if Britney Spears sang the song.
```{r}
ifelse(billboard$artist == "Spears, Britney", "yes", "no")
```

5. Test your function on the billboard data. Save the results to a new column called "wasItBritney"
```{r}
billboard$wasItBritney <- ifelse(billboard$artist == "Spears, Britney", "yes", "no")
cbind.data.frame(billboard, billboard$wasItBritney)
```

6. Use a for loop to determine if a song placed 50 or below or placed above 50 during the first week. 
```{r}
for(i in 1:nrow(billboard)){
  if(billboard$wk1[i] > 50){ billboard$fiftyaboveorbelow[i] <- "above 50"
  }
  if(billboard$wk1[i] <= 50){billboard$fiftyaboveorbelow[i] <- "below 50"}
}
```

7. Use an if else statement to determine the same information as you obtained with the for loop. 
```{r}
billboard$fiftyaboveorbelowpttwo <- ifelse(billboard$wk1 > 50, "above 50", "below 50")
```

8. Compare your results. 
```{r}
#View(billboard)
#they look the same! yay
```

