---
title: "ggplot part 1 homework"
author: "Cali Keller"
date: "2025-10-27"
output: html_document
---

1. open the starwars dataset in tidyverse
```{r}
library(tidyverse)
data(starwars)
```

2. create a graph showing the frequency of eye colors in the dataset. 
```{r}
ggplot(starwars, aes(eye_color))+
  geom_bar()
```


3. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc. 
```{r}
ggplot(starwars, aes(eye_color))+
  geom_bar(fill = "seagreen", color = "black")+
  theme_classic()+
  xlab("Eye Color")+
  ylab("Frequency")+
  ggtitle("Starwars Character's Eye Color Frequency")

```

4. Create a graph displaying information about character's height in each homeworld. 
```{r}
ggplot(starwars, aes(homeworld, height))+
  geom_boxplot()
```

5. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(homeworld, height))+
  geom_boxplot()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 55, hjust = 1))+
  xlab("Homeworld")+
  ylab("Height")+
  ggtitle("Starwars Character's Height vs Homeworld")

#i found the theme for the x axis on chat because i hated how ugly the lables were otherwise
```

6. Create a graph of character's height by mass. 
```{r}
ggplot(starwars, aes(mass, height))+
  geom_point()
```


7.Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(mass, height))+
  geom_point(color = "lightblue")+
  xlab("Mass")+
  ylab("Height")+
  theme_classic()+
  ggtitle("Starwars Character's Height by Mass")
  
```

8. Add gender as a third variable to your graph of height by mass. 
```{r}
ggplot(starwars, aes(mass, height, color = gender))+
  geom_point()
```

9. Customize your graph to make it more presentable. Consider color, backgrounds, axes, etc.
```{r}
ggplot(starwars, aes(mass, height, color = gender))+
  geom_point()+
  xlab("Mass")+
  ylab("Height")+
  scale_color_manual(name = "Gender", values = c("pink", "lightblue"))+
  ggtitle("Starwars Characters Mass and Height by Gender")+
  theme_classic()+
  facet_wrap(~gender)

```

