---
title: "ggplot pt 2 assignment"
author: "Cali Keller"
date: "2025-11-04"
output: html_document
---
---
title: "ggplot part 2 assignment"
author: "Amanda Mae Woodward"
date: "2025-10-24"
output: html_document
---

1. Open the starwars dataset (located in tidyverse)
```{r}
library(tidyverse)
data(starwars)
```

2. Create a scatter plot of character's height and mass. 
```{r}
library(ggplot2)
ggplot(starwars, aes(height, mass))+
  geom_point()
```

2a. Change the color and shape of the points in your scatter plot. 
```{r}
library(ggplot2)
ggplot(starwars, aes(height, mass))+
  geom_point(color = "blue", pch = 15)
```

3. Add Homeworld as a third variable to your scatterplot. 
```{r}
ggplot(starwars, aes(height, mass, fill = homeworld))+
  geom_point()
```

3a. What do you notice about the graph? (e.g. describe what you see)
```{r}
#the graph has many different worlds and is really ugly. theres one very clear outlier but thats it
```

4. Make a boxplot of height. 
```{r}
ggplot(starwars, aes(height))+
  geom_boxplot()
```

5. Make a boxplot of height for each species present in the data. 
```{r}
ggplot(starwars, aes(height, fill = species))+
  geom_boxplot()
```

5a. change the appearance of your graph in at least 3 ways. 
```{r}
ggplot(starwars, aes(height, fill = species))+
  geom_boxplot(color = "brown", size = .2)+
  theme_classic()
```

6. Using any variables in the starwars dataset, create a graph that uses a color palette from viridis.
```{r}
library(viridis)
ggplot(starwars, aes(height, fill =sex))+
  geom_histogram()+
  scale_fill_viridis(discrete = TRUE, option = "A")
```

7.Using any variables in the starwars dataset, create a graph that uses a color palette from RColorBrewer.
```{r}
library(RColorBrewer)
ggplot(starwars, aes(height, mass, color = sex))+
  geom_point()+
  scale_color_brewer(palette = "GnBu")
```
8. Using any variables in the starwars dataset, create a graph using facet_wrap
```{r}
ggplot(starwars, aes(mass))+
  geom_histogram()+
  facet_wrap(~sex)
```

