---
title: "psy 4802 quiz 1"
author: "Cali Keller"
date: "2025-10-13"
output: html_document
---
#don't need both headings

This quiz will cover everything we've covered in foundations of R assignments.

1. Open the tidyverse library.
```{r}
library(tidyverse)
```
√
2. Open the starwars dataset.
```{r}
data("starwars")
```
√
3. Create a subset of data that contains only the humans from the starwars franchise.
```{r}
humansonly <- subset(starwars, species == "Human")
```
√
4. In this subset of data, who is the tallest character? 
```{r}
#Darth Vader
```
√
5. In this subset of data, print the information about the character in the third row. 
```{r}
humansonly[3,]
```
√
6. In the full starwars dataset, calculate the average mass for the masculine characters. 
```{r}
masculine <- filter(starwars, gender == "masculine")
mean(masculine$mass, na.rm = TRUE)
```
√
7. Create a dataset that contains only the characters and the films that they appeared in from the full dataset. 
```{r}
pplandfilms <- starwars %>% select(name, films)
```
√
8. Create a new column in the starwars dataset that contains a "ranking" of all characters (Note: you don't have to actually rank them. Just fill the column in with the numbers 1-87)
```{r}
starwars$ranking <- 1:87
```
√
9. Create a subset of data for the characters from Naboo. How many characters are from Naboo?
```{r}
naboo <- subset(starwars, homeworld == "Naboo")
#11 characters are from naboo!
```
√
10. How many droids are in the full dataset? 
```{r}
droid <- starwars$species == "Droid"
summary(droid)
#there are 6 droids in the data set
```
√
