---
title: "statsinRassignment"
author: "Cali Keller"
date: "2025-11-18"
output: html_document
---

1. Open the diamonds dataset from ggplot2.   
```{r}
library(ggplot2)
data(diamonds)
```
2. Run a correlation between the x and y variables. What do you notice?
```{r}
cor(diamonds$x, diamonds$y)
#they are highly correlatted
```

3. What is the correlation between price and carat?
```{r}
cor(diamonds$price, diamonds$carat)
```

4. Create a linear model predicting price from table. What is the coefficient for table? 
```{r}
lm(price~table, data = diamonds)
```

5. Run a regression and an Anova predicting price from color. What do you notice about the two types of output? 
```{r}
a1 <- lm(price ~ color, data = diamonds)
anova(a1)
```

6. Run a t test to determine if "D" color diamonds are more expensive than "E" color diamonds. 
```{r}
D <- diamonds$color == "D"
E <- diamonds$color == "E"
t.test(D, E)
```

7. Open the warpbreaks dataset. What test would you use to determine the correlation between wool type and breaks?
```{r}
data(warpbreaks)
```

8. What is the correlation between wool type and breaks? 
```{r}
levels(warpbreaks$wool) <- c(0,1)
warpbreaks$wool <- as.numeric(levels(warpbreaks$wool)[warpbreaks$wool])
cor(warpbreaks$wool, warpbreaks$breaks)
```

9. A student has a 35 question, multiple choice test. Each question has 4 choices. If they answered 17 questions correctly, did their responses differ from chance?
```{r}
binom.test(17, 35, .25)
#yes, that is higher than by chance.
```

